How quantum mechanics requires non-additive measures

### 2023-10-22 Proceedings/2023-QIP/QuantumMeasure.tex
QIP: reviewed

The fact that quantum theory is a non-classical probability theory is well established. Quantum probability cannot be recovered assuming the standard axioms of Kolmogorov style probability, and their standard implementation in terms of measure theory. Various alternatives have been developed with different goals: some intend to give a more general account of probability, where classical and quantum cases are subcases; some want to offer better tools for calculation; some are developed to answer more interpretative or philosophical questions. Arguably the most well-known and used alternative approach is the use the quasi-probability distribution over phase space, either in the Wigner or the Glauber–Sudarshan P representation.

While the unsuitability of standard measure theory to represent quantum probability is a crucial aspect, this focus seems to have obscured the fact that measure theory plays another crucial in physical theories: it is used to ``count', or better to ``quantify'', states. As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which aims to start from the laws and find a minimal set of physical assumptions needed to rederive the laws. What we found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution. The former (i.e. the symplectic structure) is the only structure that allows to quantify states, and configurations over independent degrees of freedom, in a way that is observer independent. If one imposes deterministic and reversible evolution, that structure has to be preserved in time, leading to Hamiltonian evolution (i.e. a symplectomorphism). All elements of classical mechanics, including canonical variables, the Lagrangian and the action principle, can be rederived and understood in terms of quantifying the flow of states. Given the formal similarity between classical and quantum mechanics, we believe a similar approach is possible for the latter. Which raises the question: what is the quantum analogue of the Louisville measure, the measure that quantifies states in the classical theory?

Statistical mechanics and information theory mandate a link between entropy and state quantification, as the Shannon/Gibbs entropy over uniform distributions corresponds to the logarithm of the count of states. In the quantum case, we can start with the von Neumann entropy and define a measure that has the same connection. What we find is that this measure is non-additive and that its features can help us better understand what quantization is, why it requires contextuality and the failure of classical probability, and why the latter can be recovered separately within each context.

The goal of this article is to propose and articulate the goals of this new line of research. It provides the vision for future works on both the mathematical details and the conceptual structure. In section II we will show how measure theory is used to quantify elements in a set and how this is needed to assign probability and probability densities to each element. In section II we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Louisville measure. We will find that this measure is not additive.\footnote{Technically, measures are defined to be additive, therefore what we find is not a measure, but a set function. However, this term is less widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).} In section IV we will see that this non-additivity is linked to non-contextuality. In section V we will see that quantization means putting a lower unitary bound on the quantification of states, while keeping a finite value over finite regions. We will see that this very directly requires non-additivity and therefore non-contextuality. In section VI we will argue that quantization of space-time means using a quantized measure on the quantification of degrees of freedom. Lastly, in section VII we will outline a series of goals for a research program that aims to fully characterized the quantized measure.

@@

Much of the focus in quantum foundations lies in the fact that quantum mechanics does not follow the rules of classical Kolmogorov probability. Given that measure theory is typically used to formalize classical probability, numerous efforts start to find an appropriate generalization of measure theory that...The investigation of the use measure theory has been predominantly tied to the Given that this are typically expressed as

However, measure theory is not used only to characterize probability. More in general, measures are used to determine the size of sets. Even in probability theory, a probability measure assigns probability to sets, and to associate probability to points one needs two measures: the probability measure $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$. Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ given us the average probability per unit size. As we shrink $U$ to a single point $x$, we will find the probability associated to that point. Since we are interested in physical theories, the measure $\mu(U)$ quantifying the number of states within the region $U$.

@@

In the continuous case, things are a bit different. If $X$ is a manifold, a set of local coordinates induce a Lebesgue measure, which can be used to quantify the number of possible value combinations. Note that a single state has measure zero, unlike the discrete case. The Lebesgue measure is defined by two properties: countable additivity and that the size of parallelepipeds corresponds to the volume expressed by the product of the difference of coordinates. Naturally, this is a problem in physics given that the quantification of the number of states should be the same for all observers.

What most miss, is that the structure of classical phase space (i.e. symplectic structure) is exactly the structure that allows to define a quantifying measure for states that is invariant under change of observers.(TODO cite book, two phil papers) Given a change of space-time variables $q^\alpha$, conjugate momentum change covariantly, which means volumes of phase space are invariant. The role of quantifying measure in classical mechanics, then, is assigned to the Louisville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e. canonical coordinates). The amount of probability for each state is given by the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density.

Note that, in both cases, the quantifying measure is potentially unbounded, as the state space can have potentially infinitely many states, as it is the case for one spatial degree of freedom. Therefore the quantifying measures are implicitly different from probability measures, and they require specific attention.

While quantum mechanics does not obey classical probability, we do use of classical probability measure and quantifying measures in some cases. Measurement outcomes are described in terms of classical probability. If we have a discrete observable, like the energy level of a harmonic oscillator, we would use the counting measure; if we have a continuous observable, like position, we would use the Lebesgue measure associated to the variable. Preparations are also in terms of classical probability, and we also use both types of quantifying measures. For example, the maximally mixed state for a qubit can be written as $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a uniform mixture over all possible states $\int_\mathcal{H} \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.

Note how, in both cases, the quantifying measure is selected by the type of preparation or measurement, even though the state space is the same. But this means that these quantifying measure cannot be possibly quantifying the number of states as they would give contradicting results. This makes the problem rather crisp: the space of preparations and measurement outcomes are both classical, yet they cannot be connected by classical means. That is, we cannot simply put a classical probability of transition between the two because we are possibly connecting rather different quantifying measures.

To sum up, the typical foundational question is: how do we generalize probability measures for quantum mechanics? We are asking something different: how do we generalize measures that quantify the number of states in quantum mechanics?

@@

While the standard structure of quantum mechanics does not have an answer to our previous question, statistical mechanics and quantum information theory have an implicit one. Statistical mechanics tells us that the thermodynamic entropy of a system, which is a physically meaningful quantity that has numerous empirical consequences, corresponds to the Shannon entropy for a classical discrete system, the Gibbs entropy\footnote{By Gibbs entropy we mean the differential Shannon entropy $-\int_X \rho \log \rho d\mu$ calculated over phase space using position and momentum (i.e. canonical coordinates).} for a classical continuous system, and the von Neumann entropy for a quantum system.

There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as we will be always measured in bits or nats. Physically, the Boltzmann constant is needed to define temperature in Kelvin. tatistical mechanics can be equivalently formulated in terms of beta $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Louisville volume in the continuous case. The two expressions for entropy are linked as the Shannon/Gibbs entropy over a uniform distribution matches the logarithm of the measure of the support. That is, if $\rho_U$ is a uniform distribution over $U$, then $S(\rho_U) = \log \mu(U)$.

In quantum mechanics, as we said before, we do not have a measure to quantify states. There is, however, a somewhat related expression: the maximum entropy attainable by a quantum system described by a Hilbert space $\mathcal{H}$ is the dimensionality $\dim_{\mathbb{C}}(\mathcal{H})$ of the space.

A similar connections can be established in information theory. The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see TODO: cite our paper.} This, again, is a quantity that leads to empirical consequences. If we have $n$ bits at our disposal, we have $2^n$ possible combination, possible messages, that can be encoded. Therefore the logarithm of the number of possible messages corresponds to the entropy. Each different message can be encoded into a different state, therefore $\log \mu(U)$ tells us the number of bits that can be encoded using the states in that region. This will exactly correspond to the number of bits required by a source characterized by a uniform probability distribution over that number of symbols.

@@

The simplest case is when $U = \{ \psi \}$ is a single state. In this case, we have $S(|\psi\>\<\psi|) = 0$ and therefore $\mu({\psi}) = 2^{0} = 1$. Note that this is the same as the classical discrete case.

The next case is when $U = \{ \psi, \phi \}$ is the set of two states. A uniform distribution over two states will correspond to the mixed state
\begin{equation}
	\rho = \frac{1}{2} |\psi\>\<\psi| + \frac{1}{2} |\phi\>\<\phi|.
\end{equation}
The entropy of said state will depend on the probability $p=|\<\psi | \phi \>|^2$ in the following way
\begin{equation}
	S(\rho) = - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} 
	- \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2}.
\end{equation}
If the states are orthogonal, then $p=0$, and $S(\rho) = 1$ therefore $\mu(U) = 2^1 = 2$. In this case we do have that $\mu(\{ \psi, \phi \}) = \mu(\{\psi\}) + \mu(\{\phi\})$, and therefore $\mu$ is additive. We can also see that if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \log |U|$ is the logarithm of the cardinality of $U$. Therefore the quantifying measure $\mu(U) = |U|$ becomes the counting measure. But this is a special case that corresponds to the highest entropy reachable with a mixture. In all other cases the entropy will be lower, and therefore, in general, $\mu(\{ \psi, \phi \}) \leq \mu(\{\psi\}) + \mu(\{\phi\})$. In other words, the measure $\mu$ is not additive.

To sum up, the physically motivated way to quantify the number of states in quantum mechanics is not additive. When counting states in quantum mechanics, $1+1 \leq 2$.

@@

While unusual claims are common in quantum mechanics, we want to have a clear, physically tenable, reason as to why the quantifying measure for quantum states must be non-additive. In the context of a spin 1/2 system, for example, why is it that $|z^+\>$ and $|z^-\>$ count as two states while $|z^+\>$ and $|x^+\>$ count as less than two?

@@

While contextuality is often presented as a somewhat abstract, almost metaphysical, property of quantum mechanics, we have a rather more down-to-earth view. The selection of the measurement means a difference in physical process of either the preparation or measurement device. Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer or magnetic field. Something is changed in the physical world, not just on the system, but around the system, on its environment. If Amanda claims that a particle is in the $|z^+\>$ state, then Boris is able to infer something about the preparation/measuring device.\footnote{This is not at all particular to quantum mechanics. We can study the motion of a cannonball here on earth, but not on the surface of the sun, as the cannonball would be immediately vaporized. The mere claim that a system exists implies that the boundary conditions necessary for that system to exist are satisfies. In this sense, all physics is contextual.} Therefore we can understand the context simply as those conditions on the environment that are necessary on the system to be found in that particular state. A set of orthogonal states, then, has the property that they can be found in the same external conditions.

Given that we have, in general, a correlation between system and environment, it is natural to understand why the contextuality leads to a non-additive measure. Two states, in fact, should be counted as two only if they lead to a change in the system. That is, \textbf{states should be quantified all-else-being-equal}. While two states from different contexts are indeed two distinct states, stating we transitioned from one to the other does not give us information only about the system: it tells us something about the environment as well. The degree of incompatibility of observables, then, ultimately measure the degree of change when realizing the context physically.

@@

Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Louisville measure $\mu_c$ for the classical continuous case and the newly defined quantum measure $\mu_d$ for the quantum case. This requires a space where all three can be defined, which fortunately exists: the surface of a sphere. First, we want to compare the behavior on a single point. The counting measure returns one, and so does the quantum measure. The Louisville measure, however, returns zero. Second, we want to compare the behavior on a finite region, an open set. The Louisville measure returns the solid angle, which is always going to be finite as it must be less than or equal to $4\pi$. The von Neuman entropy of a mixed state over a two state system is at most one, and therefore the measure must be less than or equal to $2$. However, given that an open set on a sphere has infinitely many point, the counting measure over a finite region will be infinite. Lastly, we want to compare the additivity. Both classical measures are additive, while the quantum measure is not.

@@

So, in the light of this, what is quantization? If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Louisville (or Lebesgue) measure will work. But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense. Low value of the quantifying measure correspond to low values of entropy and measure one corresponds do zero entropy. Regions with measure less than one correspond to negative thermodynamic entropy, which also does not make sense. In fact, classical statistical mechanics does fail in those regimes: the classical equations of state for an ideal gas to predict negative entropy, which is fixed by the Fermi and Bose statistics.

@@

In our view, this provides good conceptual reasons to understand the failure of classical mechanics, the need for the departure from classical ideas and the need for quantum theory. We believe this is already sufficient reason to start a research program to explore these ideas fully. However, since most physicists are not interested in this type of ``clean-up work'', let's see this is, in our mind, critical for the development of future physical theories.

@@

As we showed before, this line of reasoning gives us a good conceptual understanding of what quantization is: the quantifying measure for states must have a unitary lower bound. If we were to quantize space-time, what is it that we are quantizing?

The most fundamental theories that are available to us are field theories. In a field theory, the state is described by the value of the fields at each point. At each point, the value of the field is independent from the others: the degrees of freedom form a continuum.\footnote{If we restrict ourselves to continuous fields, then we only need to specify the value of the fields on a countable dense subset (i.e. rational values of the coordinates). Though these kind of details do not change the discussion: a Lebesgue measure on the rationals have the same characteristics of the one extended to the reals.} We can therefore pose the question: how do we quantify the degrees of freedom?

We have again the same problem as before: measures on the continuum are not invariant under generic coordinate transformations. We do have a sense, however, that if we double the spatial volume, we double the number of degrees of freedom, and spatial volume is an invariant. Therefore we can use a measure of volume constructed from the metric tensor. This would be the equivalent of the Louisville measure we got from phase space, and we end up with an equivalent problem.

Since the measure is finite for finite regions and is additive, as we shrink the region, we will find regions with less than one degree of freedom. This does not make sense. If we want to quantize space, then, we need to put a unitary lower bound to the measure. But now our measure cannot be additive: what does it mean that two degrees of freedom count less than two?

Before answering that question, let's add time. We said that if we double the spatial volume, we double the number of degrees of freedom. Suppose we have a space-time region and we double the time, are we doubling the number of degrees of freedom? Well, we certainly doubling the support of the function we have to specify for the field configuration, but we wouldn't consider that doubling the degrees of freedom. Intuitively, we just think of those as being the same degrees of freedom that are getting different values. This is because we expect the future and past values to be linked by some dynamical equations, so the future values would be linked, in some way, to the past values. That is, the past and future value of the same field at some point are not independent.

So, what does it mean that two degrees of freedom count less than two? It could simply mean that they are not independent. That is, the values of the field at two distant points can be chosen arbitrarily, but as the points get sufficiently close, this would not be the case. This would make sense. Note that the lack of independence of the field values in close-by regions may mean that high frequencies of the fields are damped: the values have to become closer. The quantized quantifying measure of the degrees of freedom would effectively provide a cutoff. This may provide a natural high energy cutoff for quantum field theories.

In our view, the arguments laid here highlight an undeniable problem and seem to offer a natural step to take in line with the change from classical to quantum theory. While the development of a quantum theory of gravity is not our primary interest, the mathematical tools needed to address the above problem are of the same nature as the one we propose to develop a non-additive measure theory to quantify quantum states.

@@

Having outlined the scope and motivation for this new approach, we now outline the main questions we think need to be addressed.

Both the counting measure and the Lebesgue/Louisville measures are unique in some sense. For the counting measure, once we define the value for a single point, typically one, the whole measure is determined. Similarly, once we have defined a real variable, the measure is determined. The Louisville requires a bit more structure, but once the canonical coordinates are chosen, the measure is determined as well. Therefore the first question is whether these quantized measure are also unique in some sense.

While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measure, that is measure that increase as the set increase. Unfortunately, the quantized measure does not have this feature. Consider the set $\{ z^+, z^- \}$. An equal mixture will return the maximally mixed state which, as we said, would correspond to a measure equal to 2. Now consider the set $\{ z^+, z^-, x^+ \}$. This will not correspond to the maximally mixed state, and therefore the measure will be less than two. The set is bigger, but the measure is smaller. Therefore we may need to develop an appropriate theory.

>>>>

It is well-established that standard measure theoretic tools cannot be used to characterize quantum probability as quantum theory is a non-classical (i.e. non-Kolmogorov) probability theory. To obviate this problem, various alternatives and extensions to measure theory and probability theory have been developed with different goals. Our interest in measure theory, however, is not in its use to capture classical probability. Measures play another crucial use in physical theories: ``counting'', or better ``quantifying'', states. It is this use that will be the focus of this work.

As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive the laws. We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution. The former (i.e. the symplectic structure) is the only structure that allows to quantify states and configurations over independent degrees of freedom in a way that is observer independent. If one imposes deterministic and reversible evolution, state quantification must be preserved, leading to Hamiltonian evolution (i.e. a symplectomorphism). All elements of classical mechanics, including canonical variables, the Lagrangian and the action principle, can be rederived and understood in terms of quantifying the flow of states. Given the formal similarity between classical and quantum mechanics, we believe a similar approach is possible for the latter. Which raises the question: what is the quantum analogue of the Louisville measure, the measure that quantifies states in the classical theory?

Statistical mechanics and information theory mandate a link between entropy and state quantification, as the Shannon/Gibbs entropy over uniform distributions corresponds to the logarithm of the count of states. In the quantum case, we can start with the von Neumann entropy and define a measure that has the same connection. What we find is that this measure is non-additive and that its features can help us better understand what quantization is, why it requires both contextuality and the failure of classical probability, and why the latter can be recovered separately within each context. Therefore we believe studying this object is useful not only for the above stated goals of reverse physics, but more in general to get a better understanding of the foundations of quantum mechanics.

The goal of this article is to propose and articulate the goals of this new line of research. It provides the overall picture in which future works will add further mathematical detail and physical insights. In section II we will show how measure theory is used to quantify elements in a set and how this is needed to assign probability and probability densities to each element. In section III we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Louisville measure. We will find that this measure is not additive.\footnote{Technically, measures are defined to be additive, therefore what we find is not a measure, but a set function. However, this term is not widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).} In section IV we will see that this non-additivity is linked to non-contextuality. In section V we will see that quantization means putting a lower unitary bound on the quantification of states, while keeping a finite value over finite regions. We will see that this very directly requires non-additivity and therefore non-contextuality. In section VI we will argue that quantization of space-time means using a quantized measure on the quantification of degrees of freedom. Lastly, in section VII we will outline a series of goals for a research program that aims to fully characterized the quantized measure.

@@

As we said in the introduction, measure theory is not only used to characterize Kolmogorovian probability. More in general, measures are used to define the size of sets. Even in probability theory, since probability measures assigns probability to sets of point, to associate probability to points one needs two measures: the \textbf{probability measure} $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.\footnote{Measure assigns a value to elements to a $\sigma$-algebra that can be thought as a set of experimental procedures.} Since we are interested in physical theories, the measure $\mu$ quantifies the number of states. Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ given us the average probability per unit size. As we shrink $U$ to a single point $x$, we will find the probability associated to that point, and how this is done depends on the nature of $\mu$.

@@

In the continuous case, things are a bit different.\footnote{Note that ``discrete' and ``continuous'' have different meaning depending on context. At a set theoretic level, continuum is associated with the cardinality of a set. In topology, discrete and continuous spaces are characterized by  properties of open sets, not cardinality: a manifold can be given a discrete topology. In the context of quantifying measures in physics, discrete means that we can count elements with natural numbers, while continuous means we use real valued quantities with units, such as meters for distances, to define the size of regions.} If $X$ is a manifold, a set of local coordinates induce a Lebesgue measure, which can be used to quantify the number of possible value combinations. Note that a single state has measure zero, unlike the discrete case. The Lebesgue measure is defined by two properties: countable additivity and that the size of parallelepipeds corresponds to the volume expressed by the product of the difference of coordinates. Naturally, this is a problem in physics given that the quantification of the number of states should be the same for all observers.

However, the structure of classical phase space (i.e. symplectic structure) is exactly the structure that allows to define a quantifying measure for states that is invariant under change of observers.(TODO cite book, two phil papers) The role of quantifying measure in classical mechanics, then, is assigned to the Louisville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e. canonical coordinates).\footnote{The volume of an infinitesimal parallelepiped is given by $dq^1 dq^2 \cdots dp_1 dp_2$. Since $dq^i$ is controvariant and $dp_i$ is covariant under change of reference, the volume remains the same.} The amount of probability for each state is given by the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density, not a the probability itself.

%Note that, in both cases, the quantifying measure is potentially unbounded, as the state space can have potentially infinitely many states, as it is the case for one spatial degree of freedom. Therefore the quantifying measures are implicitly different from probability measures, and they require specific attention.

Quantum mechanics, instead, does not provide a clear state quantifying measure. When we characterizing preparations and measurement outcomes, we may use the discrete or continuous quantifying measures in different cases. For example, consider a particle. If we are interested in a discrete observable, like the energy level of a harmonic oscillator, we would use the counting measure; if we are interested in a continuous observable, like position, we would use the Lebesgue measure associated to the variable. As for preparations, the same maximally mixed state for a qubit can be written as discrete mixture $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a continuous uniform mixture over all possible states $\int_\mathcal{H} \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.

To be precise, the quantifying measures we considered are really on preparations and measurements, not the states themselves. It was the discrete or continuous character of the space of preparations that selected the appropriate space with the appropriate measure. But we can have different preparations and measurement spaces, the state space of system should always be the same, with always the same way to count states! This, in other words, highlights the problem of quantum probabilities: while the space of preparations and measurement outcomes are both classical, they cannot be connected by classical means. That is, we cannot simply put a classical probability of transition between the two because we are possibly connecting rather different quantifying measures.

To sum up, the typical foundational question is: how do we generalize probability measures for quantum mechanics? We are asking something different: given that state-quantifying measures play an equally fundamental role, how do we quantify the number of states in quantum mechanics?

@@

While the standard structure of quantum mechanics does not give us an answer to our question, statistical mechanics and quantum information theory have an implicit answer. Statistical mechanics tells us that the thermodynamic entropy of a system, which is a physically meaningful quantity that has numerous empirical consequences, corresponds to the Shannon entropy for a classical discrete system, the Gibbs entropy\footnote{By Gibbs entropy we mean the differential Shannon entropy $-\int \rho \log \rho d\mu$ calculated over phase space using position and momentum (i.e. canonical coordinates).} for a classical continuous system, and the von Neumann entropy for a quantum system.

There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as we will be always measured in bits or nats. Physically, the Boltzmann constant is needed to define temperature in Kelvin. Statistical mechanics can be equivalently formulated in terms of beta $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Louisville volume in the continuous case. The two expressions for entropy are linked as the Shannon/Gibbs entropy over a uniform distribution matches the logarithm of the measure of the support. That is, if $\rho_U$ is a uniform distribution over $U$, then $S(\rho_U) = \log \mu(U)$. In quantum mechanics, as we said before, we do not have a measure to quantify states. There is, however, a somewhat related expression: the maximum entropy attainable by a quantum system described by a Hilbert space $\mathcal{H}$ is the dimensionality $\dim_{\mathbb{C}}(\mathcal{H})$ of the space. Therefore, $ S(\rho_U) \leq \dim_{\mathbb{C}}(\text{span}(U))$

A similar connection can be established in information theory. The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see TODO: cite our paper.} This, again, is a quantity that leads to empirical consequences. If we have $n$ bits at our disposal, we have $2^n$ possible combination, possible messages, that can be encoded. Therefore the logarithm of the number of available messages corresponds to the number of bits that can be encoded using those messages. Physically, we can use different states to encode different messages, therefore $\log \mu(U)$ tells us the number of bits that can be encoded using the states in that region. This will exactly correspond to the number of bits required by a source characterized by a uniform probability distribution over that number of symbols.

@@

We can start by calculating the simplest case, which is when $U = \{ \psi \}$ is a single state. In this case, we have $S(|\psi\>\<\psi|) = 0$ and therefore $\mu(\{\psi\}) = 2^{0} = 1$. The measure returns one for each state as in the classical discrete case.

The next case is when $U = \{ \psi, \phi \}$ is the set of two states. A uniform distribution over two states will correspond to the mixed state
\begin{equation}
	\rho = \frac{1}{2} |\psi\>\<\psi| + \frac{1}{2} |\phi\>\<\phi|.
\end{equation}
The entropy of said state will depend on the probability $p=|\<\psi | \phi \>|^2$ in the following way
\begin{equation}
	S(\rho) = - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} 
	- \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2}.
\end{equation}
If the states are orthogonal, then $p=0$, and $S(\rho) = 1$ therefore $\mu(U) = 2^1 = 2$. In this case we have that $2 = \mu(\{ \psi, \phi \}) = \mu(\{\psi\}) + \mu(\{\phi\}) = 1 + 1$, and therefore $\mu$ is additive. Moreover, if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \dim_{\mathbb{C}}(\text{span}(U)) = \log |U|$ is the logarithm of the cardinality of $U$. Therefore the quantifying measure $\mu(U) = |U|$ is equal to the counting measure. But this is a special case that corresponds to the highest entropy reachable with a mixture. In all other cases the entropy will be lower, and therefore, in general, $\mu(\{ \psi, \phi \}) \leq \mu(\{\psi\}) + \mu(\{\phi\})$. In other words, the measure $\mu$ is not additive.

To sum up, the physically motivated way to quantify the number of states in quantum mechanics is, in general, not additive. When counting states in quantum mechanics, $1+1 \leq 2$.

@@

While unusual claims are common in quantum mechanics, we want to have a clear, physically tenable, reason as to why the quantifying measure for quantum states must be non-additive. In the context of a spin 1/2 system, for example, why is it that $\{ |z^+\>, |z^-\>\}$ count as two states while $\{ |z^+\>, |x^+\> \}$ count as less than two?

@@

While contextuality is often presented as a somewhat abstract, almost metaphysical, property of quantum mechanics, we have a rather more down-to-earth view. The selection of the measurement means a difference in physical process of either the preparation or measurement device. Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer, magnetic field or equivalent change to a physical device. Something is changed in the physical world, not just on the system, but around the system, on its environment. If Amanda claims that a particle is in the $|z^+\>$ state, then Boris is able to infer something about Amanda's preparation/measuring device.\footnote{This is not at all particular to quantum mechanics. We can study the motion of a cannonball here on earth, but not on the surface of the sun, as the cannonball would be immediately vaporized. The mere claim that a system exists implies that the boundary conditions necessary for that system to exist are satisfies. In this sense, all physics is contextual.} Therefore we can understand the context simply as those conditions on the environment that are necessary on the system to be found in that particular state. A set of orthogonal states, then, has the property that they can be found in the same external conditions.

Given that we have, in general, a correlation between system and environment, it is natural to understand why the contextuality leads to a non-additive measure. Two states, in fact, should be counted as two only if they lead to a change in the system alone. That is, \textbf{states should be quantified all-else-being-equal}. While two states from different contexts are indeed two distinct states, stating we transitioned from one to the other does not give us information only about the system: it tells us something about the environment as well. The degree of incompatibility of observables, then, ultimately measure the degree of change when realizing the context physically.

@@

Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Louisville measure $\mu_c$ for the classical continuous case and the newly defined quantized measure $\mu_q$ for the quantum case. This requires a space where all three can be defined, which fortunately exists: the surface of a sphere. First, we want to compare the behavior on a single point. The counting measure returns one, and so does the quantum measure. The Louisville measure, however, returns zero. Second, we want to compare the behavior on a finite region, an open set. The Louisville measure returns the solid angle, which is always going to be finite as it must be less than or equal to $4\pi$. The von Neuman entropy of a mixed state over a two state system is between zero and one, and therefore the measure must be between one and $2$. However, given that an open set on a sphere has infinitely many point, the counting measure over a finite region will be infinite. Lastly, we want to compare the additivity. Both classical measures are additive, while the quantum measure is not.

@@

So, in the light of this, what is quantization? If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Louisville (or Lebesgue) measure will work. But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense. Low value of the quantifying measure correspond to low values of entropy and measure one corresponds do zero entropy. Regions with measure less than one correspond to negative thermodynamic entropy, which also does not make sense. In fact, classical statistical mechanics does fail in those regimes: the classical equations of state for an ideal gas to predict negative entropy, which is fixed by the Fermi and Bose statistics.

@@

In our view, this provides good conceptual reasons to understand the failure of classical mechanics, the need for the departure from classical ideas and the need for quantum theory. We believe this, by itself, is already sufficient reason to start a research program to explore these ideas fully. However, since most physicists do not seem to be interested in this type of ``clean-up work'', we can easily argue that a full characterization of the quantized measure is critical for the development of future physical theories.

@@

We have seen that quantization can be understood as putting a lower unitary bound on the state quanfication: no set of states can have fever than one state. If we were to quantize space-time, what is it that we are quantizing?

The most fundamental theories that are available to us are field theories. In a field theory, the state is described by the value of the fields at each point in space. At each point, the value of the field is independent from the others: the degrees of freedom form a continuum.\footnote{If we restrict ourselves to continuous fields, then we only need to specify the value of the fields on a countable dense subset (i.e. rational values of the coordinates). Though these kind of details do not change the discussion: a Lebesgue measure on the rationals have the same characteristics of the one extended to the reals.} We can therefore pose the question: how do we quantify the degrees of freedom?

If we have a region of space and we double the volume, we can imagine we are also doubling the number of degrees of freedom. Therefore, we can construct a volume measure from the metric tensor, which would give us an invariant volume that quantifies degrees of freedom, much in the same way that the Louisville measure quantified states.

Since the measure is finite for finite regions and is additive, as we shrink the region, we will find regions with less than one degree of freedom. This does not make sense. As before, we need to put a unitary lower bound to the measure and our measure cannot be additive. We need a quantized measure. Quantizing space-time, means putting a quantized measure on the number of degrees of freedom.

What does it mean that two degrees of freedom count less than two? Before answering that question, let's consider time. We said that if we double the spatial volume, we double the number of degrees of freedom. Suppose we have a space-time region and we double the time, are we doubling the number of degrees of freedom? Well, we certainly doubling the support of the function we have to specify for the field configuration, but we wouldn't consider that doubling the degrees of freedom. Intuitively, we just think of those as being the same degrees of freedom that are getting different values. This is because we expect the future and past values to be linked by some dynamical equations, so the future values would be linked, in some way, to the past values. That is, the past and future value of the same field at some point are not independent.

The same field at two points, then, may count less than to degrees of freedom simply because they are not independent. The values of the field at two distant points can be chosen arbitrarily, therefore the measure would be additive at long range. But as the points get sufficiently close, this would not be the case, so at short scale we need a quantized measure to quantify independent degrees of freedom. Note that the lack of independence of the field values in close-by regions may mean that high frequencies of the fields are damped: the values have to become closer. A quantized measure of the degrees of freedom would effectively provide a cutoff. Therefore we have a story that is very well motivated and provide a possible solution for high energy divergences in quantum field theories.

While the development of a quantum theory of gravity is not our primary interest, the mathematical tools required to characterize the non-additive quantized measure seem to be of similar nature as the ones that may be used to quantize space-time.

@@

Having outlined the scope and motivation for this new approach, we now outline the main questions we think need to be addressed by this new research program.

Both the counting measure and the Lebesgue/Louisville measures are unique in some sense. For the counting measure, once we define the value for a single point, typically one, the whole measure is determined. Similarly, once we have defined a real variable, the Lebesgue measure is determined. The Louisville requires a bit more structure, but once the canonical coordinates are chosen, the measure is determined as well. Therefore the first question is whether these quantized measure are also unique in some sense.

While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increase. Unfortunately, the quantized measure does not have this feature. Consider the set $\{ z^+, z^- \}$. An equal mixture will return the maximally mixed state which, as we said, would correspond to a measure equal to 2. Now consider the set $\{ z^+, z^-, x^+ \}$. This will not correspond to the maximally mixed state, and therefore the measure will be less than two. The set is bigger, but the measure is smaller. Therefore there may be new mathematical ground to be explored.
